\documentclass{article}

\usepackage{aistats2018}
 % If your paper is accepted, change the options for the package
% aistats2018 as follows:
%
%\usepackage[accepted]{aistats2018}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

\input{commands}

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{Learning the effect of latent variables in Gaussian Graphical models with unobserved variables}

\aistatsauthor{ Author 1 \And Author 2 \And  Author 3 }

\aistatsaddress{ Institution 1 \And  Institution 2 \And Institution 3 } ]


\begin{abstract}
Gaussian graphical models (GGM) have been widely used in many high-dimensional applications. Problems in which all relevant variables have been identified is extremely rare. The unobserved variables (or latent variables) can induce misleading dependence structure between observed variables. In this paper, we focus on a family of Latent Variable Gaussian graphical models (LVGGM), where the model is conditionally sparse given latent variables, but marginally non-sparse. In order to reveal latent variables and recover the structure of the complete model, we introduce a convex formulation with a new regularization imposing sparsity on latent factors.  We propose a tractable convex algorithm and study identifiability conditions. We show promising results on synthetic \TODO{and real datasets (remove)}.\\
\end{abstract}



\input{intro}
 
\input{related}

\input{setup}

\input{identifiability}

\input{experiments}

\subsection*{Conclusion}
We consider a family of latent variable Gaussian graphical model whose concentration matrix has a sparse plus low-rank structure. We introduce a regularization to impose more structure on the low rank component, as a sum of multiple low-rank and $k$-sparse factors, and derive a convex formulation. We propose an efficient algorithm to optimize this problem and show promising results for recovering the structure of the complete GGM in the experiments. We provide a first result for identifiability condition. Futurework includes extending the framework to multiple levels of sparsity and generalizing identifiability conditions for recovery guarantees on our optimisation problem .
%
%\subsubsection*{Acknowledgements}
%
%Use unnumbered third level headings for the acknowledgements.  All
%acknowledgements go at the end of the paper.

%\subsubsection*{References}
\bibliographystyle{apalike}
\bibliography{lvggm}

\end{document}
