\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts,amsthm}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{MnSymbol}
\usepackage{color}
\usepackage{graphicx} % more modern
\usepackage{subfigure} 
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{lineno}
\usepackage{thmtools,thm-restate}
\usepackage{xr}

%\linenumbers



\def\A{\mathcal{A}}
\def\ga{\gamma_\mathcal{A}}
\def\Co{C^{\circ}}
\def\Coa{C^{\circ}_{\!\!\mathcal{A}}}
\def\Ca{C_{\!\mathcal{A}}}
\def\go{\gamma^{\circ}}
\def\goa{\gamma^{\circ}_{\A}}
\newcommand{\dt}[2]{\langle#1,#2\rangle}
\def\Olasso{\Omega_{\rm Lasso}} % Lasso norm
\def\Ogl{\Omega_{\rm GL}} % group Lasso norm
\def\Olgl{\Omega_{\rm LGL}} % latent group Lasso norm
\def\tr{{\rm tr}}
\newcommand\itgset[1]{[\!\![#1]\!\!]}
\def\RR{\mathbb{R}}
\def\EE{\mathbb{E}}
\newcommand{\tblue}[1]{\textcolor{blue}{#1}}
\newcommand{\tred}[1]{\textcolor{red}{#1}}
\newcommand\TODO[1]{\tblue{TODO: \texttt{#1}}}
%\newcommand\TODO[1]{}
\newcommand\OLD[1]{\tblue{#1}}
\newcommand\NEW[1]{\tred{#1}}
\def\st{\text{s.t.}}
\def\supp{\text{supp}}

\def\BIT{\begin{itemize}}
\def\EIT{\end{itemize}}
\def\BET{\begin{enumerate}}
\def\EET{\end{enumerate}}
\def\ra{$\:\rightarrow\:$}
\def\y{\mathbf{y}}
\def\x{\mathbf{x}}
\def\v{\mathbf{v}}
\def\u{\mathbf{u}}
\def\w{\mathbf{w}}
\def\s{\mathbf{s}}
\def\g{\mathbf{g}}
\def\h{\mathbf{h}}
\def\X{\mathbf{X}}
\def\RR{\mathbb{R}}
\def\ei{\varepsilon_i}
\def\ej{\varepsilon_j}
\def\T{\mathcal{T}}
\def\I{\mathcal{I}}

\def\op{{\rm op}}
\def\supp{{\rm Supp}}
\def\st{\text{s.t.}}
\newcommand\normop[1]{|||#1|||_{\infty}}
\newcommand\normbop[1]{|\!|\!|#1|\!|\!|_{\infty,2}}
\newcommand\normopop[1]{|\!|\!|#1|\!|\!|_{\rm op,\rm op}}
\newcommand\ve[1]{\varepsilon^{{\scriptscriptstyle#1}}}
\newcommand\sss[1]{{\scriptscriptstyle#1}}
\def\uI{U^{\sss{I}}}
\def\xxi{\zeta}

\newtheorem{thm}{Theorem}
\newtheorem{prop}{Proposition}
\newtheorem{fact}{Fact}
\newtheorem{lemm}{Lemma}
\newtheorem{coro}{Corollary}
\newtheorem{mydef}{Definition}

\title{Learning the effect of latent variables in Gaussian Graphical models with unobserved variables}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  David S.~Hippocampus\thanks{Use footnote for providing further
    information about author (webpage, alternative
    address)---\emph{not} for acknowledging funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
Gaussian graphical models (GGM) have been widely used in many high-dimensional applications. Some effects can affect the data, that is, some of the relevant variables may be latent. In this paper, we focus on a family of Latent Variable Gaussian graphical models (LVGGM), where the model is conditionally sparse given latent variables, but marginally non-sparse. In order to identify the effect of each latent variable and recover the structure of the complete model, we introduce a convex formulation with a new regularization imposing more structure on latent factors.  We propose a tractable convex algorithm and study identifiability conditions. We show promising results on synthetic and real datasets.\\
\end{abstract}



\input{intro}
 
\input{related}

\input{setup}

\input{identifiability}

\input{experiments}

\subsection*{Conclusion}
We consider a family of latent variable Gaussian graphical model whose concentration matrix has a sparse plus low-rank structure. We introduce a regularization to impose more structure on the low rank component, as a sum of multiple low-rank and $k$-sparse factors, and derive a convex formulation. We propose an efficient algorithm to optimize this problem and show promising results for recovering the structure of the complete GGM in the experiments. We provide a first result for identifiability condition. Futurework includes extending the framework to multiple levels of sparsity and generalizing identifiability conditions for recovery guarantees on our optimisation problem .


%\section*{References}
\bibliographystyle{apalike}
\bibliography{lvggm}

\end{document}
